---
title: "STAT 327 Homework 3"
output: html_document
---

We'll grade your homework by

* opening your "hw3.Rmd" file in RStudio
* clicking "Knit HTML"
* reading the HTML output
* reading your "hw3.Rmd"

You should write R code anywhere you see an empty R code chunk. You
should write English text anywhere you see "..."; please surround it
with doubled asterisks (`**...**`) so that it will show up as boldface
and be easy for us to find.

Include reasonable labels (titles, axis labels, legends, etc.) with
each of your graphs.

Name: Elaine Huang

Email: xhuang266@wisc.edu

We'll use data on housing values in suburbs of Boston. They are in an
R package called "MASS." (An R package is a collection of code,
data, and documentation. "MASS" refers to the book "Modern Applied
Statistics with S." R developed from the earlier language, S.) The
MASS package comes with the default R installation, so it's already on
your computer.  However, it's not loaded into your R session by
default. So we'll load it via the `require()` command (there's nothing
for you to do here):

```{r}
require("MASS")
```

Run `?Boston` (outside this R Markdown document) to read the help page for the `Boston` data frame.

Convert the `chas` variable to a factor with labels "off" and "on" (referring to the Charles river).

```{r}
Boston$chas = factor(Boston$chas, levels = c("0" , "1"), labels = c("off", "on"))
Boston$chas
```

How many rows are in the Boston data frame? How many columns?

```{r}
dim(Boston)
n.rows = dim(Boston)[1]
n.cols = dim(Boston)[2]
cat(sep = "", "rows = ", n.rows, "\n")
cat(sep = "", "columns = ", n.cols, "\n")
```

What does a row represent?
`**Each row represents a unique town in Boston area.**`

What does a column represent?
`**Each column represents a different statistic that is being measured for each town.**`

Make a density plot (with rug) of tax rates.

```{r}
plot(density(Boston$tax), main = "Full-value property-tax rate per $10,000 in Boston")
rug(Boston$tax)
```

Describe the shape of the distribution of tax rates.
`**The shape of the distribution of tax rates is a bimodal distribution. The probability density function would possess two "humps"; one from 0-506 with the peak around 300, and another one after 506 with the peak around 700.**`

Note that the distribution shape doesn't make sense in light of the rug representation of the data. Make a histogram of the tax rates.

```{r}
hist(Boston$tax, freq = TRUE, main = "Histogram of Boston tax rate", xlim = c(0, 800), ylim = c(0, 150))

```

Why is the second peak of the density plot so large? In what way is the rug representation of the data inadequate? Write a line or two of code to figure it out, and then explain it.

```{r}
max = sum(Boston$tax == max(Boston$tax))
cat(sep = "", "The maximum value ", max(Boston$tax) , " occurs " , max, " times in the graph.", "\n")
second.max = sort(Boston$tax, decreasing = TRUE)[max + 1]
secondmax = sum(Boston$tax == second.max)
cat(sep = "", "The second maximum value ", second.max, " occurs ", secondmax, " times in the graph", "\n")
```
`**Since the maximum value 711 occurs 5 times in the graph while the second maximum value occurs 132 times, which has a high frequency, the second peak of the density plot is very large. Even though the biggest value occurs five times and the second biggest value occurs 132 times, the rug representation only shows one on the graph, which inadequately represents the frequency.**`

Make a barplot of "chas".

```{r}
barplot(table(Boston$chas), main = "Barplot of Charles River Counts", ylim = c(0, 500))
```

How many neighborhoods are on the Charles river?

```{r}
river.count = length(which(Boston$chas == "on"))
cat(sep = "", "There are ", river.count, " neighborhoods on the Charles river.", "\n")
```

Make a single graph consisting of three plots: 

* a scatterplot of "nox" on the y-axis vs. "dis" on the x-axis 
* a (vertical) boxplot of "nox" left of the scatterplot's y-axis
* a (horizontal) boxplot of "dis" below the scatterplot's x-axis

Hint: use `layout()` with a 4x4 matrix, using the top-right 3x3 corner
for the scatterplot, leaving the bottom-left 1x1 corner blank, and
using the other parts for the boxplots.

(An optional challenge, worth 0 extra credit points: remove the axis
and plot border from each boxplot.)

```{r, fig.width=7, fig.height=7}

a = matrix(data = c(2,1,1,1,2,1,1,1,2,1,1,1,0,3,3,3), nrow = 4, ncol = 4, byrow = TRUE)
layout(a)
plot(Boston$dis, Boston$nox, xlab = "Boston$dis", ylab = "Boston$nox", main = "Nitrogen Oxides Concentration vs. Weighted mean of distances" )
boxplot(Boston$nox, horizontal = FALSE)
boxplot(Boston$dis, horizontal = TRUE)


```

Look into the highest-crime neighborhood by making a single graph of one column of three rows:

* Find the row number, r, of the neighborhood with the highest "crim".
* Make a density plot of "crim". Include a rug to show the data.
* Add a red circle at (x, y) = (max crime rate, 0) to make this maximum crime rate stand out.
* Make a density plot with rug of "medv", adding a red circle at (x, y) = (medv[r], 0) to see what medv corresponds to the highest crime rate.
* Repeat the last step for "ptratio".

```{r}
m = matrix(data = c(1,2,3),nrow = 3, ncol = 1, byrow = TRUE)
layout(m)

n = order(Boston$crim)
r = n[length(n)]

plot(density(Boston$crim),  main = "Boston per capita crime rate by town")
rug(Boston$crim)
points(x = max(Boston$crim), y =0, col= "red", type = "p", pch = 1, cex = 2)


plot(density(Boston$medv), main = "Median value of owner-occupied homes in Boston")
rug(Boston$medv)
points(x = Boston$medv[r], y =0, col= "red", type = "p", pch = 1, cex = 2)

plot(density(Boston$ptratio), main = "Pupil-teacher ratio by town in Boston", xlim = c(10, 25))
rug(Boston$ptratio)
points(x = Boston$ptratio[r],y =0, col= "red", type = "p", pch = 1, cex =2)






```

What do you notice about the ptratio and medv for the highest-crime neighborhood?

`**In the highest-crime neighborhood, pupil-teacher ratio is around 20 and median value of owner-occupied homes in \$1000s is around 10.
Thus, for the highest-crime neighborhood,the pupil-teacher ratio is relatively high and median value of owner-occupied homes in \$1000s is relatively low. **`